I offer an earnest appreciation for the various professional communities that contributed to the development of this thesis:
\begin{itemize}
\item[]  \textbf{Abundant Intelligences} project contextualizing Artificial Intelligence within Indigenous knowledgeways, with important feedback from Archer Pechawis and Peter Morin

\item[] \textbf{American Society for Cybernetics} as Transgenerational Collaborator to its president, Dr. Paul Pangaro

\item[] \textbf{Creative Reaction Labs} as a student with important feedback from founder Antionette D. Carroll

\item[] \textbf{Digital Futures graduate program} at OCAD U as a graduate student, with important feedback from Dr. Emma Westecott, Kate Hartman, and Cindy Poremba

\item[] \textbf{Global Centre for Climate Action} as graduate student member, with important feedback from its Associate Directors Dr. Ian Clarke and Nick Puckett

\item[] \textbf{Harvard Divinity School} as a presenter, exhibitor, and member of various working groups under Dr. Dan McKanan in the Program for the Evolution of Spirituality, and in collaboration with Dr. Eric Mortensen, Zia Pollis, Owen Yager, and Sadie Trichler

\item[] \textbf{Institute for Equitable Design and Justice} as a founding member 

\item[] \textbf{Interdisciplinary Master's in Art, Media, and Design graduate program} at OCAD U, the program in which I began my degree, as an exhibitor and exhibition organizer, with important feedback from Graduate Program Director Jay Irizawa

\item[] \textbf{Perceptual Artifacts Lab} for cross-sensory accessibility research at OCAD U, as lab contributor and presenter, with important feedback from its director, Dr. Peter Coppin

\item[] \textbf{Strategic Foresight and Innovation graduate program} at OCAD U, as a cross-department participant and presenter for in its 2024 International Innovation Forum, with important feedback from Graduate Program Director Suzanne Stein

\item[] \textbf{Strategic Innovation Lab} as a research associate, with important feedback from its co-founders Greg Van Alstyne and Dr. Peter Jones

\item[] \textbf{Systemic Design Association} as a member and reviewer for the thirteenth Relating Systems Thinking and Design Symposia (RSD13), with important feedback from Dr. Birger Sevaldson, Dr. Evan Timothy Barba, Ryan J. A. Murphy, and Cheryl May

\item[] \textbf{Visual Analytics Lab} as a research assistant with important feedback from its director, Dr. Sara Diamond
\end{itemize}