\section*{Mentors and Advisors}
To my mentors, for keeping me safe and well, especially Dr. Monisha Bhatia, Dr. Allison Lou, Christelle Thibeault, Germán Sanmateu, Vit\sout{u}nk\sout{u}y Herrera Ubate, Alex McLaren, Dr. Robyn Stephens, Erlene Webster, Roxie Ateah, Dr. Sat Chuen Hon, and Richard Kwan.

\section*{Students}
To my students, whose enthusiasm has fuelled this work, from OCAD University, the Amity School of Engineering \& Technology, and Collective 65. I treasure our shared journey of discovery.

\section*{Co-learners}
In OCAD University's graduate programs—Digital Futures, Strategic Foresight and Innovation, Interdisciplinary Master's of Art, Media and Design, and Criticism and Curatorial Practice— I found support and friendship with Abha, Afifa, Aishwarya, Anas, Anusha, Aranya, Benny, Cait, Camille, Casper, David, Dedra, Deepikah, Elfy, Elsha, E.J., Flavia, Gayatri, Jennifer-Amy, Jessica, Jiyeon, Juan, Kunal, Lo, Maryam, Matthew, Mona, Mufaro, Nasim, Nastaran, Nikhil, Prayag, Purvi, Raha, Raquel, Rhea, Ricardo, Rim, Roderick, Sal, Sally, Samarth, Sara, Shannyn, Shipra, Sissie, Spoorti, Stephanie, Susan, Tallulah, Tamika, Taylor, Termeh, Tyler, Valen, Yiqun, Yitong, Youssef, Yu, Zaheen, and Zubin. I note Alberto Castillo and Hala Alsalman with appreciation for their wit, insight, and hospitality, which was meaningful to me and many others.


\section*{Friends}
Thank you to my dear friends for inspiring me daily, especially Adam, Anne, Brittan, Caro, Caleb, David, Douglas, Elena, Gordon, Jason, Jessica, Jesús, Juan Pablo, Justin, Kristin, Leanne, Leonardo, Masha, Michael, Nicholas, Nilabjo, Oyin, Pippa, Sanja, Sheree, Stellan, Stephanie, Vanessa, Warren, and Willem.

\section*{Family}
To the ancestors whose stories my family retells as a source of inspiration, hope, and resilience, especially Mario Castaño Pérez and Mariela Castaño Pérez, for their enduring examples of fortitude and solidarity.

To my mother Gloria Margarita Castaño Suárez, my first teacher of design, and matriarch of my family, for sustaining our stories across the expanses of continents; to my father Carlos Arturo Castaño, for being our family’s brave advocate and enduring sustainer while we faced seemingly impossible odds; to my elder brother Carlos Agusto, for your intercession; to my younger brother Simón, for your mindful rebelliousnes and exuberant support, and for being my hope. To Angela, for bringing joy to my brother. To my partner Carmen, for lighting up my home with tender companionship.

\section*{The More-Than-Human World}
To the many more-than-human beings I am in relationship with in my life. I name with special warmth the waters, especially Lake Ontario, for calling me and insisting that I visit. As an extension of my parents, sitting with you makes it easier for the harmony which exists before language to resurface. To the red-tailed hawks, thank you for visiting on days I was in most need of re-encountering trust in the invisible parts of life.