\section*{Supervisors and Advisors}
Central to this work's development was the radically interdisciplinary support of my primary supervisor Dr. Adam Tindale from OCAD U Digital Future's graduate program. I also offer thanks to my secondary supervisor Greg van Alstyne from the OCAD U Strategic Foresight and Innovation program, and to my external reviewer Ali Shamas Qadeer.

\section*{OCAD University}
My sincere thanks go to the OCAD U educators and advisors who shaped my work through their support and feedback. While many have contributed to my research, I particularly acknowledge support from: Peter Morin, Dr. Cindy Poremba, Kate Hartman, Simone Jones, Dr. Emma Westecott, Suzanne Stein, Dr. Amish Morrell, Dr. Dot Tuer, Dr. Michael J. Prokopow, Dr. Ashok Mathur, Dr. Michelle Wyndham-West, Lillian Allen, Alda Escareño, Angela Bains, Dr. Kathy Moscou, Isabel Meirelles, Natalie Majaba Waldburger, Michael Lee Poy, Dr. Kathy Kiloh, Jay Irizawa, Dr. Alexis Morris, Dr. Ian Clarke, Dr. Fidelia Lam, Dr. Julian Haladyn, Dr. Neal Thomas, Dr. Suzanne Morrissette, Dr. Barbara Rauch, Peter Scott, Michelle Gay, JJ Lee, Dr. Peter Coppin, Dr. Nabil Harfoush, Dr. Sara Diamond, Dr. Susan Blight, Reagan Kennedy, Melissa General, Aleena Derohanian, Shreya Chopra, Sandy Soud, Omar Qureshi, Meichen Waxer, Martin Chochinov, Kimberly Gwanzura, Hannah Dickson, Elizabeth Clydesdale, and Gale Allen.

\section*{Harvard University}
I offer my earnest appreciation to my colleagues, friends, and mentors from the Harvard Divinity School Program for the Evolution of Spirituality and its various working groups. I extend my gratitude in particular to Dr. Dan McKanan, Dr. Eric Mortensen, Zia Pollis, Owen Yager, and Sadie Trichler. Thank you for your calls to convene in the liminal spaces of many ways of knowing.

\section*{External Institutions}
My research has been enriched by collaborations with numerous institutions. I acknowledge with professional admiration contributors to this work's development from: American Society for Cybernetics, City University of New York, Concordia University, European Council for Nuclear Research/Conseil Européen pour la Recherche Nucléaire (CERN), Georgetown University, COMSTATS University Islamabad, Indian Institute of Technology, Institute of Equitable Design and Justice and Creative Reaction Labs, Mason Studio, Moebio Labs, New York University (NYU), Nodus Labs, Oslo School of Architecture and Design, Redress Movement, Swarthmore College, Systemic Design Association, TerminusDB, University of Waterloo, University of British Columbia, University of Florida, University of Newfoundland, and the University of Toronto.

\section*{Editors and Publishing Support}
The drafting of this thesis owes much to the dedication of Dr. Inder Sidhu, Marsha Taichman, Chris Landry, Bryan Weissenboeck, Rebecca Diederichs, Dr. Lennart Nacke, Douglas Davey, and Abdullah Askar. Their editorial insights and technical expertise spanned from conceptual clarity to LaTeX implementation, and helped shape this thesis.